\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-11-13 tex/gitlinks.tex
more text

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/150c31cb799f4eb003031cfb856cd3b53ab5f5fc/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/150c31cb799f4eb003031cfb856cd3b53ab5f5fc/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/150c31cb799f4eb003031cfb856cd3b53ab5f5fc/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

>>>>

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/bea234c69e977f48e7acb13bb803f991e89aa4e9/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/bea234c69e977f48e7acb13bb803f991e89aa4e9/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/bea234c69e977f48e7acb13bb803f991e89aa4e9/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-13 tex/ms.tex
more text

\section{Spherical Harmonics}
\label{sec:ylm}
%
The spherical harmonics $Y_{l,m}$ are a complete set of orthogonal functions on the
surface of the sphere and, as such, are a convenient basis in which to construct
our gaussian process. As in \citet{Luger2019}, here we describe the intensity
on the surface of a star as an expansion in the real spherical harmonics, which is
fully specified by a set of spherical harmonic coefficients
$\{ y_{l,m}\}$, where $l \in [0, l_{\mathrm{max}}]$ is known as
the degree and
$m \in [-l, l]$ is known as the order, with $l_{\mathrm{max}}$ equal to the highest
degree in the expansion.
It is often convenient to collect the spherical harmonic coefficients of
a given expansion into a vector $\mathbf{y}$ indexed by a single
integer $n$, where
%
\begin{align}
    \label{eq:n}
    n = l^2 + l + m
\end{align}
%
and, conversely,
%
\begin{align}
    \label{eq:lm}
    \begin{split}
        l & = \floor{\sqrt{n}}
        \\
        m & = n - l^2 - l
        \quad.
    \end{split}
\end{align}
%
\citet{Luger2019} derived the expression for the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe from a star rotating about a fixed axis
if the stellar surface intensity is well described by the spherical harmonic
coefficient vector $\mathbf{y}$:
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \pmb{\mathcal{A}} \, \mathbf{y}
    \quad,
\end{align}
%
where $\pmb{\mathcal{A}} = \pmb{\mathcal{A}}(t, i, P)$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is implicitly
a function of the time $t$, the stellar inclination $i$, and the stellar
rotation period $P$. For reference, the rows of $\pmb{\mathcal{A}}$ are given by
Equation~(18) in \citet{Luger2019}.

@@

Let
$\mathbf{f} = \left( f_0 \, f_1 \, \cdots \,  f_K \right)^\top$
denote a vector of $K$ flux measurements at times
$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_K \right)^\top$.
Conditioned on certain physical properties of the star, $\pmb{\theta}$,
we wish to compute the mean $\pmb{\mu}(\pmb{\theta})$ and
covariance $\pmb{\Sigma}(\pmb{\theta})$
of $\mathbf{f}$, which together fully specify our GP in flux.
%
As with any random variable, the mean and covariance may be computed from
the expectation value of $\mathbf{f}$ and
$\mathbf{f}\,\mathbf{f}^\top$, respectively:
%
\begin{align}
    \label{eq:mean}
    \pmb{\mu}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \pmb{\theta} \Big]
    \\
    \label{eq:cov}
    \pmb{\Sigma}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}(\pmb{\theta}) \pmb{\mu}^\top(\pmb{\theta})
    \quad.
\end{align}
%
Given the linear relationship between flux and spherical harmonic
coefficients (Equation~\ref{eq:fAy}),
we may write the mean and covariance of our flux GP as
%
\begin{align}
    \pmb{\mu}(\pmb{\theta})
     & = \pmb{\mathcal{A}}(\pmb{\theta}) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
    \\
    \pmb{\Sigma}(\pmb{\theta})
     & = \pmb{\mathcal{A}}(\pmb{\theta}) \, \pmb{\Sigma}_{\mathbf{y}} \, \pmb{\mathcal{A}}^\top(\pmb{\theta})
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:mean_y}
    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
    \\
    \label{eq:cov_y}
    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}) \pmb{\mu}_{\mathbf{y}}^\top(\pmb{\theta})
\end{align}
%
are the mean and covariance of the GP in the spherical harmonics basis.
The bulk of the math in this paper is devoted to computing
the expectations in the expressions above, which
are given by the integrals
%
\begin{align}
    \label{eq:exp_y}
    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
    \\
    \label{eq:exp_yy}
    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
    \quad,
\end{align}
%
where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
distribution of features on the surface  (i.e., the size and location of star spots)
and $p(\mathbf{x} \, \big| \, \pmb{\theta})$ is its probability density
function (PDF).

@@

and $p(\mathbf{x} \, \big| \, \pmb{\theta})$, the integrals in the expressions

@@

i.e., $p(\mathbf{x} \, \big| \, \pmb{\theta})$.

@@

For simplicity, we assume that $p(\mathbf{x} \, \big| \, \pmb{\theta})$

@@

    p(\mathbf{x} \, \big| \, \pmb{\theta})

@@

    \pmb{\theta} = \left(

@@

    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]

@@

    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big]

@@

We devote the next several sections to the computation of these eight
integrals.

\section{The Size Integral}
\label{sec:size}
%
Our goal in this section is to compute the first and second moments
of the size distribution ($\mathbf{e}_\phi$, $\mathbf{E}_\phi$) under
a suitable spherical harmonic expansion $\mathbf{s}(r)$ of the spot profile
and a suitable probability distribution function for the spot radius,
$p(r \, \big| \, \pmb{\theta}_{r})$.

\subsection{Spot profile}
%
We model the brightness $b$ an angle $\vartheta$ away from the
center of a spot of negative unit intensity and radius $r$ be
%
\begin{align}
    \label{eq:brvartheta}
    b(r; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{r-\vartheta}{s}\right)} - 1
\end{align}
%
for some (constant) shape parameter $s$. In the limit $s \rightarrow 0$, $b$ approaches an
inverted top-hat function with half-width equal to $r$,
corresponding to a circular spot of uniform intensity. For $s > 0$, each half
of $b$ is a sigmoid with half-width at half-minimum equal to $r$.
In our implementation of the algorithm we choose $s = 0.2^\circ$,
which is small compared to features of interest but not too small as to
create numerical issues when computing model gradients (which would be
undefined at the spot boundary if the spot profile were truly an inverted top-hat).

@@

\subsection{Probability density function}
%
For simplicity, we will adopt a uniform probability distribution for the
spot radius, characterized by a mean radius $r_0$ and a half-width $\Delta$:
%
\begin{align}
    p(r \, \big| \, \pmb{\theta}_{r})
     & =
    \begin{cases}
        \frac{1}{2\Delta} & r_0 - \Delta \leq r \leq r_0 + \Delta
        \\
        0                 & \mathrm{otherwise}
        \quad,
    \end{cases}
\end{align}
%
where the hyperparameters of the distribution are
%
\begin{align}
    \pmb{\theta}_r = \left(
    r_0 \, \, \, \,
    \Delta \right)^\top
    \quad.
\end{align}
%
As we argue in the text, in practice it is often difficult to constrain the
moments of the radius distribution above the first (the mean). It is
therefore useful to also consider the limiting case of the radius distribution
as $\Delta \rightarrow 0$, in which case the PDF becomes
%
\begin{align}
    p(r \, \big| \, \pmb{\theta}_{r}, \Delta = 0)
     & =
    \delta(r - r_0)
    \quad,
\end{align}
%
where $\delta$ is the delta function.
%

\subsection{First moment}
%
The first moment of the radius distribution is (Equation~\ref{eq:e1})
%
\begin{align}
    \mathbf{e}_r
     & \equiv
    \int
    \mathbf{s}(r) \,
    p(r \, \big| \, \pmb{\theta}_{r}) \,
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{1}{2\Delta}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    \mathbf{s}(r) \,
    \mathrm{d}r
    \quad.
\end{align}
%
Using the equations from the previous section, its components may be written
%
\begin{align}
    (e_r)^l_m
     & =
    \frac{\delta_{m,0}}{2\Delta}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    \tilde{s}_{l}(r) \,
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{\delta_{m,0}}{2\Delta}
    \sum_{k=0}^{K-1} B^+_{l,k}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    b_{k}(r)
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{\delta_{m,0} s}{2\Delta}
    \sum_{k=0}^{K-1} B^+_{l,k}
    \ln
    \left(
    \frac{
        1+\exp\left(\frac{r -\Delta -\vartheta_k}{s}\right)
    }
    {
        1+\exp\left(\frac{r + \Delta -\vartheta_k}{s}\right)
    }
    \right)
    \quad.
\end{align}
%
In the limit $\Delta \rightarrow 0$,
%
\begin{align}
    \lim_{\Delta \rightarrow 0}
    \mathbf{e}_r
     & =
    \mathbf{B}^+ \mathbf{b}(r)
    \quad.
\end{align}
%

\subsection{Second moment}
%
The second moment of the radius distribution is (Equation~\ref{eq:E1})
%
\begin{align}
    \mathbf{E}_r
     & \equiv
    \int
    \mathbf{s}(r) \,
    \mathbf{s}^\top(r) \,
    p(r \, \big| \, \pmb{\theta}_{r}) \,
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{1}{2\Delta}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    \mathbf{s}(r) \,
    \mathbf{s}^\top(r) \,
    \mathrm{d}r
    \quad.
\end{align}
%
As before, its components may be written
%
%
\begin{align}
    (E_r)^{l,l'}_{m,m'}
     & =
    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    \tilde{s}^l_{m}(r) \,
    \tilde{s}^{l'}_{m'}(r) \,
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
    \sum_{k=0}^{K-1} B^+_{l,k}
    \sum_{k'=0}^{K-1} B^+_{l',k'}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    b_{k}(r)
    b_{k'}(r)
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{\delta_{m,0}\delta_{m',0}s}{2\Delta}
    \sum_{k=0}^{K-1} B^+_{l,k}
    \sum_{k'=0}^{K-1} B^+_{l',k'}
    C_{k,k'}(r)
    \quad,
\end{align}
%
where
%
\begin{align}
    C_{k,k'}(r)
     & \equiv
    \frac{
        \exp\left(\frac{\vartheta_{k} - \vartheta_{k'}}{s}\right)
        \ln
        \left(
        \frac{
            1+\exp\left(\frac{r -\Delta -\vartheta_{k}}{s}\right)
        }
        {
            1+\exp\left(\frac{r + \Delta -\vartheta_{k}}{s}\right)
        }
        \right)
        -
        \ln
        \left(
        \frac{
            1+\exp\left(\frac{r -\Delta -\vartheta_{k'}}{s}\right)
        }
        {
            1+\exp\left(\frac{r + \Delta -\vartheta_{k'}}{s}\right)
        }
        \right)
    }{
        1
        -
        \exp\left(\frac{\vartheta_{k} - \vartheta_{k'}}{s}\right)
    }
    \quad.
\end{align}
%
In the limit $\Delta \rightarrow 0$,
%
\begin{align}
    \lim_{\Delta \rightarrow 0}
    \mathbf{E}_r
     & =
    \mathbf{B}^+ \mathbf{b}(r) \mathbf{b}^\top(r) {\mathbf{B}^+}^\top
    \quad.
\end{align}
%

\section{The Latitude and Longitude Integrals}
\label{sec:lat-lon}
%
Our goal in this section is to compute the first and second moments
of the latitude ($\mathbf{e}_\phi$, $\mathbf{E}_\phi$)
and longitude ($\mathbf{e}_\lambda$, $\mathbf{E}_\lambda$) distributions
(Equations \ref{eq:e2}, \ref{eq:e3}, \ref{eq:E2}, and \ref{eq:E3},
respectively). These involve integrals over the terms in the Wigner
rotation matrix for spherical harmonics, which we discuss below.

>>>>

\section{Spherical Harmonics}
\label{sec:ylm}
%
The spherical harmonics $Y_{l,m}$ are a complete set of orthogonal functions on the
surface of the sphere and, as such, are a convenient basis in which to construct
our gaussian process. As in \citet{Luger2019}, here we describe the intensity
on the surface of a star as an expansion in the real spherical harmonics, which is
fully specified by a set of spherical harmonic coefficients
$\{ y_{l,m}\}$, where $l \in [0, l_{\mathrm{max}}]$ is known as
the degree and
$m \in [-l, l]$ is known as the order, with $l_{\mathrm{max}}$ equal to the highest
degree in the expansion.
It is often convenient to collect the spherical harmonic coefficients of
a given expansion into a vector $\mathbf{y}$ indexed by a single
integer $n$, where
%
\begin{align}
    \label{eq:n}
    n = l^2 + l + m
\end{align}
%
and, conversely,
%
\begin{align}
    \label{eq:lm}
    \begin{split}
        l & = \floor{\sqrt{n}}
        \\
        m & = n - l^2 - l
        \quad.
    \end{split}
\end{align}
%
\citet{Luger2019} derived the expression for the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe from a star rotating about a fixed axis
if the stellar surface intensity is well described by the spherical harmonic
coefficient vector $\mathbf{y}$:
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \pmb{\mathcal{A}} \, \mathbf{y}
    \quad,
\end{align}
%
where $\pmb{\mathcal{A}}$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is implicitly
a function of the stellar inclination $i$, the stellar
rotation period $P$, the stellar limb darkening coefficients, etc.
For reference, the rows of $\pmb{\mathcal{A}}$ are given by
Equation~(18) in \citet{Luger2019}.

@@

Let
$\mathbf{f} = \left( f_0 \, f_1 \, \cdots \,  f_K \right)^\top$
denote a vector of $K$ flux measurements at times
$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_K \right)^\top$.
Conditioned on certain physical properties of the
star
(such as its inclination, rotation period, etc.), $\pmb{\theta}_\star$,
and on certain properties of the  star spots
(such as their sizes, positions, etc.), $\pmb{\theta}_\bullet$,
we wish to compute the mean $\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ and
covariance $\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$
of $\mathbf{f}$, which together fully specify our GP in flux.
%
As with any random variable, the mean and covariance may be computed from
the expectation value of $\mathbf{f}$ and
$\mathbf{f}\,\mathbf{f}^\top$, respectively:
%
\begin{align}
    \label{eq:mean}
    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big]
    \\
    \label{eq:cov}
    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet) \pmb{\mu}^\top(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
    \quad.
\end{align}
%
Given the linear relationship between flux and spherical harmonic
coefficients (Equation~\ref{eq:fAy}),
we may write the mean and covariance of our flux GP as
%
\begin{align}
    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
    \\
    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(\pmb{\theta_\star})
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:mean_y}
    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
    \\
    \label{eq:cov_y}
    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] - \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet) \pmb{\mu}_{\mathbf{y}}^\top(\pmb{\theta}_\bullet)
\end{align}
%
are the mean and covariance of the GP in the spherical harmonics basis.
The bulk of the math in this paper is devoted to computing
the expectations in the expressions above, which
are given by the integrals
%
\begin{align}
    \label{eq:exp_y}
    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
    \\
    \label{eq:exp_yy}
    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
    \quad,
\end{align}
%
where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
distribution of features on the surface  (i.e., the size and location of star spots)
and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability density
function (PDF).

@@

and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$, the integrals in the expressions

@@

i.e., $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$.

@@

For simplicity, we assume that $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$

@@

    p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)

@@

    \pmb{\theta}_\bullet = \left(

@@

    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]

@@

    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]

@@

We devote the next section to the computation of these eight
integrals.

\section{The Expectation Integrals}
\label{sec:integrals}
\subsection{The Size Integral}
\label{sec:size}
%
Our goal in this section is to compute the first and second moments
of the size distribution ($\mathbf{e}_\phi$, $\mathbf{E}_\phi$) under
a suitable spherical harmonic expansion $\mathbf{s}(r)$ of the spot profile
and a suitable probability distribution function for the spot radius,
$p(r \, \big| \, \pmb{\theta}_{r})$.

\subsubsection{Spot profile}
%
We model the brightness $b$ an angle $\vartheta$ away from the
center of a spot of negative unit intensity and radius $r$ be
%
\begin{align}
    \label{eq:brvartheta}
    b(r; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{r-\vartheta}{s}\right)} - 1
\end{align}
%
for some (constant) shape parameter $s$. In the limit $s \rightarrow 0$, $b$ approaches an
inverted top-hat function with half-width equal to $r$,
corresponding to a circular spot of uniform intensity. For $s > 0$, each half
of $b$ is a sigmoid with half-width at half-minimum equal to $r$.
In our implementation of the algorithm we choose $s = 0.2^\circ$,
which is small compared to features of interest but not too small as to
create numerical issues when computing model gradients (which would be
undefined at the spot boundary if the spot profile were truly an inverted top-hat).

@@

\subsubsection{Probability density function}
%
For simplicity, we will adopt a uniform probability distribution for the
spot radius, characterized by a mean radius $r_0$ and a half-width $\Delta$:
%
\begin{align}
    p(r \, \big| \, \pmb{\theta}_{r})
     & =
    \begin{cases}
        \frac{1}{2\Delta} & r_0 - \Delta \leq r \leq r_0 + \Delta
        \\
        0                 & \mathrm{otherwise}
        \quad,
    \end{cases}
\end{align}
%
where the hyperparameters of the distribution are
%
\begin{align}
    \pmb{\theta}_r = \left(
    r_0 \, \, \, \,
    \Delta \right)^\top
    \quad.
\end{align}
%
As we argue in the text, in practice it is often difficult to constrain the
moments of the radius distribution above the first (the mean). It is
therefore useful to also consider the limiting case of the radius distribution
as $\Delta \rightarrow 0$, in which case the PDF becomes
%
\begin{align}
    p(r \, \big| \, \pmb{\theta}_{r}, \Delta = 0)
     & =
    \delta(r - r_0)
    \quad,
\end{align}
%
where $\delta$ is the delta function.
%

\subsubsection{First moment}
%
The first moment of the radius distribution is (Equation~\ref{eq:e1})
%
\begin{align}
    \mathbf{e}_r
     & \equiv
    \int
    \mathbf{s}(r) \,
    p(r \, \big| \, \pmb{\theta}_{r}) \,
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{1}{2\Delta}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    \mathbf{s}(r) \,
    \mathrm{d}r
    \quad.
\end{align}
%
Using the equations from the previous section, its components may be written
%
\begin{align}
    (e_r)^l_m
     & =
    \frac{\delta_{m,0}}{2\Delta}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    \tilde{s}_{l}(r) \,
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{\delta_{m,0}}{2\Delta}
    \sum_{k=0}^{K-1} B^+_{l,k}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    b_{k}(r)
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{\delta_{m,0} s}{2\Delta}
    \sum_{k=0}^{K-1} B^+_{l,k}
    \ln
    \left(
    \frac{
        1+\exp\left(\frac{r -\Delta -\vartheta_k}{s}\right)
    }
    {
        1+\exp\left(\frac{r + \Delta -\vartheta_k}{s}\right)
    }
    \right)
    \quad.
\end{align}
%
In the limit $\Delta \rightarrow 0$,
%
\begin{align}
    \lim_{\Delta \rightarrow 0}
    \mathbf{e}_r
     & =
    \mathbf{B}^+ \mathbf{b}(r)
    \quad.
\end{align}
%

\subsubsection{Second moment}
%
The second moment of the radius distribution is (Equation~\ref{eq:E1})
%
\begin{align}
    \mathbf{E}_r
     & \equiv
    \int
    \mathbf{s}(r) \,
    \mathbf{s}^\top(r) \,
    p(r \, \big| \, \pmb{\theta}_{r}) \,
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{1}{2\Delta}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    \mathbf{s}(r) \,
    \mathbf{s}^\top(r) \,
    \mathrm{d}r
    \quad.
\end{align}
%
As before, its components may be written
%
%
\begin{align}
    (E_r)^{l,l'}_{m,m'}
     & =
    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    \tilde{s}^l_{m}(r) \,
    \tilde{s}^{l'}_{m'}(r) \,
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
    \sum_{k=0}^{K-1} B^+_{l,k}
    \sum_{k'=0}^{K-1} B^+_{l',k'}
    \int_{r_0 - \Delta}^{r_0 + \Delta}
    b_{k}(r)
    b_{k'}(r)
    \mathrm{d}r
    \nonumber \\
     & =
    \frac{\delta_{m,0}\delta_{m',0}s}{2\Delta}
    \sum_{k=0}^{K-1} B^+_{l,k}
    \sum_{k'=0}^{K-1} B^+_{l',k'}
    C_{k,k'}(r)
    \quad,
\end{align}
%
where
%
\begin{align}
    C_{k,k'}(r)
     & \equiv
    \frac{
        \exp\left(\frac{\vartheta_{k} - \vartheta_{k'}}{s}\right)
        \ln
        \left(
        \frac{
            1+\exp\left(\frac{r -\Delta -\vartheta_{k}}{s}\right)
        }
        {
            1+\exp\left(\frac{r + \Delta -\vartheta_{k}}{s}\right)
        }
        \right)
        -
        \ln
        \left(
        \frac{
            1+\exp\left(\frac{r -\Delta -\vartheta_{k'}}{s}\right)
        }
        {
            1+\exp\left(\frac{r + \Delta -\vartheta_{k'}}{s}\right)
        }
        \right)
    }{
        1
        -
        \exp\left(\frac{\vartheta_{k} - \vartheta_{k'}}{s}\right)
    }
    \quad.
\end{align}
%
In the limit $\Delta \rightarrow 0$,
%
\begin{align}
    \lim_{\Delta \rightarrow 0}
    \mathbf{E}_r
     & =
    \mathbf{B}^+ \mathbf{b}(r) \mathbf{b}^\top(r) {\mathbf{B}^+}^\top
    \quad.
\end{align}
%

\subsection{The Latitude and Longitude Integrals}
\label{sec:lat-lon}
%
Our goal in this section is to compute the first and second moments
of the latitude ($\mathbf{e}_\phi$, $\mathbf{E}_\phi$)
and longitude ($\mathbf{e}_\lambda$, $\mathbf{E}_\lambda$) distributions
(Equations \ref{eq:e2}, \ref{eq:e3}, \ref{eq:E2}, and \ref{eq:E3},
respectively). These involve integrals over the terms in the Wigner
rotation matrix for spherical harmonics, which we discuss below.

\subsubsection{Rotation matrices}
%
The Wigner rotation matrix for real spherical harmonics up to degree $l_{\mathrm{max}}$
may be written as the block-diagonal matrix
%
\begin{align}
    \setstackgap{L}{1.25\baselineskip}
    \fixTABwidth{T}
    \mathbf{R} =
    \parenMatrixstack{
    \quad\quad \, \mathbf{R}^0 \, \quad\quad
     &              &              &        &                               \\
     & \mathbf{R}^1 &              &        &                               \\
     &              & \mathbf{R}^2 &        &                               \\
     &              &              & \ddots &                               \\
     &              &              &        & \mathbf{R}^{l_{\mathrm{max}}}
    }\quad,
\end{align}
%
where
%
\begin{align}
    \mathbf{R}^l = {\mathbf{U}^l}^{-1} \mathbf{D}^l \mathbf{U}^l
\end{align}
%
is the Wigner rotation matrix for a single spherical harmonic degree,
%
\begin{align}
    \label{eq:U}
    \setstackgap{L}{1.25\baselineskip}
    \fixTABwidth{T}
    \mathbf{U}^l =
    \frac{1}{\sqrt{2}}
    \parenMatrixstack{
        \quad\quad\, \ddots \, \quad\quad\quad\quad
           &       &        &       &          &    &   &    & \Ddots \\
           & \imag &        &       &          &    &   & 1  &        \\
           &       & \imag  &       &          &    & 1 &    &        \\
           &       &        & \imag &          & 1  &   &    &        \\
           &       &        &       & \sqrt{2} &    &   &    &        \\
           &       &        & \imag &          & -1 &   &    &        \\
           &       & -\imag &       &          &    & 1 &    &        \\
           & \imag &        &       &          &    &   & -1 &        \\
    \Ddots &       &        &       &          &    &   &    & \ddots
    }
\end{align}
%
describes the transformation from complex to real spherical harmonics,
and $\mathbf{D}$ is the Wigner matrix for complex spherical harmonics,
whose terms are given by the expression
%
\begin{align}
    D_{m,m'}^l(\alpha, \beta, \gamma) = \mathrm{e}^{-\imag m' \alpha}
    d_{m,m'}^l(\beta) \mathrm{e}^{-\imag m \gamma}
\end{align}
%
where $\alpha$, $\beta$, and $\gamma$ are the Euler angles describing
the rotation in the
$\hat{\mathbf{z}}{-}\hat{\mathbf{y}}{-}\hat{\mathbf{z}}$ convention
and $\imag$ is the imaginary unit. The terms of the $d$-matrix
depend on powers of $\sin\left(\nicefrac{\beta}{2}\right)$
and $\cos\left(\nicefrac{\beta}{2}\right)$
\citep[c.f. Equation C15 in][]{Luger2019}, but it is convenient to use
the half-angle formula to express these terms instead as
%
\begin{proof}{test_dlmmp}
    d_{m,m'}^l(\beta) =
    \sum\limits_{i=0}^{2l} c_{m,m',i}^{l}
    \mathrm{sgn}(\sin\beta)^{i}
    (1 - \cos\beta)^{\frac{2l - i}{2}}
    (1 + \cos\beta)^\frac{i}{2}
\end{proof}
%
where
%
\begin{proof}{test_dlmmp}
    c_{m,m',i}^{l} & =
    \resizebox{.75\hsize}{!}{$
            \begin{cases}
                \dfrac{
                \left(-1\right)^{\frac{2l + 3m + m' - i}{2}}
                \sqrt{\left(l - m\right)!
                    \left(l + m\right)!
                    \left(l - m'\right)!
                    \left(l + m'\right)!}
                }{
                2^l
                \left(\frac{i - m - m'}{2}\right)!
                \left(\frac{i + m + m'}{2}\right)!
                \left(\frac{2l - i - m + m'}{2}\right)!
                \left(\frac{2l - i + m - m'}{2}\right)!
                }
                %
                 & m - m' - i \,\, \mathrm{even}
                %
                \\[0.5em]
                0
                %
                 & m - m' - i \,\, \mathrm{odd}
            \end{cases}
        $}
\end{proof}
%

\subsubsection{Probability density function for the latitude}
%
The latitude integrals (Equations~\ref{eq:e2} and \ref{eq:E2}) involve
rotations by an angle $\phi$ about $\hat{\mathbf{x}}$, which
may be accomplished by choosing
Euler angles $\alpha = \nicefrac{\pi}{2}$, $\beta = \phi$, and
$\gamma = -\nicefrac{\pi}{2}$, such that
%
\begin{align}
    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi)
     & =
    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \mathbf{U}^l
\end{align}
%
with
\begin{align}
    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi)
     & =
    \mathbf{D}^l\left(\frac{\pi}{2}, \phi, -\frac{\pi}{2}\right)
    \quad.
\end{align}
%
From the expressions above, it is clear that all terms in
$\mathbf{R}_{\hat{\mathbf{x}}}(\phi)$ are equal to (weighted) sums of powers
of $(1 \pm \cos\phi)$.
%
Since our goal is to compute integrals of these terms multiplied by
a probability density function, it is convenient to model
$\cos\phi$ as a Beta-distributed variable. As we will see, this
choice will allow us to
analytically compute the first two moments of the distribution of
$\phi$ conditioned on $\pmb{\theta}_\phi$.
